---
title: "Data607 HW6 Part2"
author: "Ritesh Lohiya"
date: "March 10, 2018"
output: html_document
---
IS 607 - Project 2:

The goal of this assignment is to give you practice in preparing different datasets for downstream analysis work.
Your task is to:
(1) Choose any three of the "wide" datasets identified in the Week 6 Discussion items. (You may use your own dataset; please don't use my Sample Post dataset, since that was used in your Week 6 assignment!) For each of the three chosen datasets:
??? Create a .CSV file (or optionally, a MySQL database!) that includes all of the information included in the dataset. You're encouraged to use a "wide" structure similar to how the information appears in the discussion item, so that you can practice tidying and transformations as described below.
??? Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data. [Most of your grade will be based on this step!]
??? Perform the analysis requested in the discussion item.
??? Your code should be in an R Markdown file, posted to rpubs.com, and should include narrative descriptions of your data cleanup work, analysis, and conclusions.
(2) Please include in your homework submission, for each of the three chosen datasets:
??? The URL to the .Rmd file in your GitHub repository, and
??? The URL for your rpubs.com web page.

I have taken the data set about marriage rates from FiveThirtyEight's GitHub and prepared my .csv file for my Part II Project.
 
```{r}
library(knitr)
library(stringr)
library(tidyr)
library(dplyr)
library(ggplot2)
```

```{r}
marriage <- read.csv("https://raw.githubusercontent.com/Riteshlohiya/Data607-Week6/master/Marriage_rates.csv", header=TRUE, sep=",")
marriage
```

We can consider only one range for thios 25 to 34:

```{r}
marriage1 <- marriage[c(2,4:21)]
marriage1
```

Changing the column names:

```{r}
colnames(marriage1)<- c("Year", "All", "High_School", "Some_College", "Bachelor_Degree", "Bachelor_other", "Graduate_Degree", "White", "Black", "Hispanic", "New_England", "Mid_Atlantic", "Midwest", "South", "Mountain", "Pacific", "Poor", "Middle_class", "Rich")
colnames(marriage1)
head(marriage1)
```

Now use gather() to club by class:

```{r}
marriage1 <- gather(marriage1, "Class", "Percentage", 2:19)
head(marriage1)
```

Make the Percentage as numeric:

```{r}
marriage2 <- suppressWarnings(marriage1 %>% 
  mutate(Percentage = as.numeric(Percentage)))
marriage2
```

Calculate Marriage rate using Mutate():

```{r}
marriage3 <- marriage2 %>% 
  mutate(Marriage_Rate = 1 - Percentage)
head(marriage3)
```

Using Rename() for Percentage field:

```{r}
marriage4 <- dplyr::rename(marriage3, Single_Rate = Percentage)
head(marriage4)
```

Filter on Class
```{r}
marriage5 <- filter(marriage4, Class == "All")
marriage5
```

Using ggplot to understand more:

```{r}
ggplot(marriage5, aes(x = Year, y=Marriage_Rate, fill = Class)) +  
          geom_bar(stat="identity",position="dodge") + 
          xlab("Year") + ylab("Marriage_Rate") 
```

```{r}
ggplot(marriage5, aes(x = Year, y=Single_Rate, fill = Class)) +  
          geom_bar(stat="identity",position="dodge") + 
          xlab("Year") + ylab("Marriage_Rate") 
```


Marriage Rate by Year for each Class:

```{r}
marriage4$Marriage_Rate <- as.numeric(as.character(marriage4$Marriage_Rate))
ggplot(marriage4, aes(x = Year, y = Marriage_Rate, group = Class, colour = Class)) +
  geom_line() +
  geom_point() +
  theme_linedraw() +
  ggtitle("Marriage Rates by Class for Ages 25-34") +
  ylab("Marriage Rate") 
  
```

We can see that the marriage rate is decling with the passing year.

 